Cryptanalysis of SBC based post-quantum signature scheme

### 2025-06-17 rapport.tex
preuve EUF-CMA

		\begin{tcolorbox}[title=Protocol 15: Signing Algorithm, breakable, enhanced, colback=white, colframe=black]
			\textbf{Inputs:} A public key $x = (\vec{u}, \vec{v})$, a private key $w = (\vec{x}, \vec{y}) \in \text{NSBC}[\vec{u}, \vec{v}]$ and a message $m \in \{0,1\}^*$.

@@

	\begin{theorem}[\cite{feneu2023}]
		Let $\text{Hash}_0, \text{Hash}_1, \text{Hash}_m$ and $\text{Expand}$ be modeled as random oracles, and let $(N, \tau, \lambda, p)$ be parameters of the signature scheme. Let $\mathcal{A}$ be an adversary against the \textbf{EUF-CMA} security of the scheme running in time $t_{\mathcal{A}}$ and making a total of $Q_{RO}$ random oracle queries and $Q_{Sig}$ signing queries. Assuming that $F$ is an $(t_{\mathcal{A}}, \epsilon_{OWF})$-hard one-way function and that PRG is a $(t_{\mathcal{A}}, \epsilon_{PRG})$-secure psuedorandom generator then $\mathcal{A}$'s advantage in the \textbf{EUF-CMA} game is
		$$
			\epsilon_{EUF-CMA} \leq \epsilon_{OWF} + \epsilon_{PRG} + \frac{(\tau N + 2)Q^2}{2^{2\lambda}} + Pr[X+Y = \tau],
		$$
		where $Q = Q_{RO} + N_{Hash} \cdot Q_{Sig}$ with $N_{Hash} = 2 + \tau(2N - 1)$ the number of hash computations in a signature generation, and where $X, Y$ are defined as
		\begin{itemize}
			\item $X = max_{q1 \in \mathcal{Q}_1} \{X_{q1}\} \text{ with } X_{q1} \sim \mathcal{B}\left(\tau, \binom{N}{l+1}p\right)$ and
			\item $Y = max_{q2 \in \mathcal{Q}_2} \{Y_{q2}\} \text{ with } Y_{q2} \sim \mathcal{B}\left(\tau - X, 1 / \binom{N}{l}\right)$,
		\end{itemize}
		where $\mathcal{B}(n_0, p_0)$ denotes the binomial distribution with $n_0$ the number of trials and $p_0$ the success probability of each trial.
	\end{theorem}

	Proof of theorem \ref{EUF-CMA}:
	\begin{proof}
		\textbf{Il suffit de modeler le schema avec les énoncés de \cite{feneu2023} et hop on peut simplement utiliser ses preuves qui sont assez gnérales pour tout système de signature MPCitH avec une oneway function et une PRF.}

	\end{proof}

>>>>

		\begin{tcolorbox}[title=Protocol 15: Signing Algorithm, breakable, enhanced, colback=white, colframe=black]\label{Protocol_sig}
			\textbf{Inputs:} A public key $x = (\vec{u}, \vec{v})$, a private key $w = (\vec{x}, \vec{y}) \in \text{NSBC}[\vec{u}, \vec{v}]$ and a message $m \in \{0,1\}^*$.

@@

	\begin{theorem}[\cite{feneu2023}]\label{theorem_feneu}
		Let $\text{Hash}_0, \text{Hash}_1, \text{Hash}_m$ and $\text{Expand}$ be modeled as random oracles, and let $(N, \tau, \lambda, p)$ be parameters of the signature scheme. Let $\mathcal{A}$ be an adversary against the \textbf{EUF-CMA} security of the scheme running in time $t_{\mathcal{A}}$ and making a total of $Q_{RO}$ random oracle queries and $Q_{Sig}$ signing queries. Assuming that $F$ is a $(t_{\mathcal{A}}, \epsilon_{OWF})$-hard one-way function and that PRG is a $(t_{\mathcal{A}}, \epsilon_{PRG})$-secure psuedorandom generator then $\mathcal{A}$'s advantage in the \textbf{EUF-CMA} game is
		$$
			\epsilon_{EUF-CMA} \leq \epsilon_{OWF} + \epsilon_{PRG} + \frac{(\tau N + 2)Q^2}{2^{2\lambda}} + Pr[X+Y = \tau],
		$$
		where $Q = Q_{RO} + N_{Hash} \cdot Q_{Sig}$ with $N_{Hash} = 2 + \tau(2N - 1)$ the number of hash computations in a signature generation, and where $X, Y$ are defined as
		\begin{itemize}
			\item $X = max_{q1 \in \mathcal{Q}_1} \{X_{q1}\} \text{ with } X_{q1} \sim \mathcal{B}\left(\tau, \binom{N}{l+1}p\right)$ and
			\item $Y = max_{q2 \in \mathcal{Q}_2} \{Y_{q2}\} \text{ with } Y_{q2} \sim \mathcal{B}\left(\tau - X, 1 / \binom{N}{l}\right)$,
		\end{itemize}
		where $\mathcal{B}(n_0, p_0)$ denotes the binomial distribution with $n_0$ the number of trials and $p_0$ the success probability of each trial.
	\end{theorem}

	Proof of theorem \ref{EUF-CMA}:
	\begin{proof}
		We model the $SBC$ signature scheme \cite{HJ23} as a 5-round protocol (see Protocol~\ref{Protocol_sig}). By applying Theorem~\ref{theorem_feneu} from \cite{feneu2023}, and under the assumption that the underlying PRG is a $(t_{\mathcal{A}}, \epsilon_{\mathrm{PRG}})$-secure pseudorandom generator, and as established in Lemma~\ref{UB-KOA} that the $SBC$ function is a $(t_{\mathcal{A}}, \epsilon_{\mathrm{OWF}})$-secure one-way function, we conclude that the $SBC$ signature scheme achieves $EUF$-$CMA$ security.
	\end{proof}
